Generating urban morphologies at large scales

### 2019-06-03 paper.tex
.
paper_camera_ready

@@

@@

The study of forms of the built environment, and more precisely of the urban environment, has been the subject of different disciplines such as architecture, urban planning, or geography, with different approaches corresponding to various scales and processes \citep{moudon1997urban,gauthier2006mapping,kropf2009aspects}.
Establishing typologies of urban morphologies, and understanding their link with underlying urban growth processes, is nowadays a crucial issue for sustainability as a large majority of the world population live in cities and energy consumption is closely related to urban form through e.g. mobility patterns and automobile dependance \citep{newman2000sustainable}.

Although there is neither a unified definition of urban form, nor unified generative models and quantitative indicators to measure it, several approaches are close to the spirit of artificial life and generative social science~\citep{bonabeau1997classical,epstein1999agent}.
Procedural modeling \citep{watson2008procedural} aims at generating realistic cities, but is mostly focused on the visual impression given and does not consider realistic generative processes.
It is furthermore developed largely at larger scales than the one of the district \citep{Parish:2001:PMC:383259.383292}. \cite{merrell2010computer} generates in that context plans for interior of buildings, whereas \cite{cruz2017generation} uses a cellular automaton model for building morphogenesis.
Approaches linked to urban planning have focused on the spatial distribution of land-use, at multiple resolutions \citep{liu2017future}, and proposed cellular automata models for urban sprawl, generally at the scale of the metropolitan area \citep{herold2003spatiotemporal}. For example, \cite{horner2007multi} proposes a link between urban form based on land-use and commuting.

@@

The quantification of urban form is in itself covered by a vast literature.
Recent work have proposed to apply deep learning techniques directly on vector data, such as \cite{2017arXiv170902939M} does, for a worldwide classification of road networks. Such an approach avoids the question of isolating relevant features.
However, as we aim at calibrating generative models, our quantification will make more sense with interpretable measures.
%devrait on justifier pourquoi on a besoin de mesures interprétables ???
%. je pense, en reaction a la tendance deep learning etc ? (mais c'est peut etre pas pertinent en effet)
\cite{boeing2018measuring} proposes an extensive review of existing measures from a large extent of disciplines, their implications for planning and design, and the relation with urban complexity. \cite{webster1995urban} uses image processing techniques such as contrast or Fourier analysis, to extract synthetic descriptions of urban areas from satellite imaging. \cite{fumega2014identification} provide a typology of cities in relation with energy consumption in the perspective of climate change. \cite{rode2014cities} relate indicators of urban form with residential heat-energy demand. Other complexity-related approaches such as fractal dimensions have been introduced as for example by \cite{batty1987fractal}.

@@

We detail now the generative model introduced. A null model is also considered, to ensure the relevance of the morphological measures and the adjustments on these. It consists in a random grid generator, where each cell is occupied if a random uniform number between 0 and 1 is below a density parameter $d_R$.

%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{blocks}
        \caption{Blocks}
        \label{fig:blocks}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{expMixture}
        \caption{Kernel mixture}
        \label{fig:mixture}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{percolation}
        \caption{Network percolation}
        \label{fig:percolation}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{random}
        \caption{Random}
        \label{fig:random}
    \end{subfigure}
    \caption{Examples of patterns produced by the synthetic generators.}
    % TODO should provide parameters here
    \label{fig:generators}
\end{figure}
%%%%%%%%%%%%%%

@@

Simulation results and real measures are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/LGK0US}. Source code is available on the git repository of the project at \url{https://github.com/openmole/spatialdata}.
% TODO split repositories
The model and indicators were coded in scala langage for performance purposes. This furthermore allows a seamless integration into the OpenMOLE workflow engine for model exploration \citep{reuillon2013openmole}, which provides methods for numerical experiments (in our case sampling methods) and transparent access to high performance computation environments (model simulation were run on the European Grid Infrastructure for an equivalent of one year and one month CPU time).

@@

We compute the morphological indicators given above on a large sample on real urban areas. For practical computational reasons, we restrain our geographical area of study to European functional urban areas as provided by \cite{bretagnolle2019following}. We expect to already have a good representativity, although not universal, of existing urban forms with this sample, as it is known that European cities already have a significant morphological diversity \citep{le2015forme}. We collect building layouts from OpenStreetMap (illustrated in Figure~\ref{fig:osm}), as this source has been shown to have a good quality especially in Europe \citep{mooney2010towards}. Using the \texttt{osmosis} tool, buildings are filtered from the openstreetmap raw dump for Europe (downloaded from \url{http://download.geofabrik.de/} in March 2019) and inserted into a Postgis database, which can then be efficiently queried for a specific bounding box. Indeed, although the library developed provides a direct access to the OpenStreetMap API, query limitations do not allow such a systematic sampling.
%Our tool also provide a direct access to the classic and overpass API % not necessary ?
We sample $N=72,000$ points into polygons corresponding to urban areas, first by selecting the area with a uniform selection weighted by population of areas, then by drawing uniform spatial coordinates within the polygon with a polygon sampling heuristic. % TODO clarify the polygon sampler

@@

%%%%%%%%%%%%%%
\begin{figure*}
\vspace{-1.8cm}
    \centering
    (1)
    \framebox{\includegraphics[width=0.18\linewidth]{closest1_blocks}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest1_real}}\hspace{0.4cm}
    (2)
    \framebox{\includegraphics[width=0.18\linewidth]{closest2_expMixture}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest2_real}}\\\vspace{0.2cm}
    (3)
    \framebox{\includegraphics[width=0.18\linewidth]{closest3_blocks}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest3_real}}\hspace{0.4cm}
    (4)
    \framebox{\includegraphics[width=0.18\linewidth]{closest4_percolation}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest4_real}}\\
    \includegraphics[width=\linewidth]{lhscalib_projrealpcs}
    \caption{Comparison of real morphologies with patterns produced by the synthetic generators. All points are projected into the first two components computed on the real measures only. Points in dark blue correspond to real configurations. The type of generator is given by the color in legend. Dark circle correspond to the centroids of clusters on which the calibration is done, their number giving the corresponding configurations shown above. For each, we successively show the best synthetic configuration ((1) block generator with $N_B=12,m_B=5,M_B=12$, (2) exponential mixture generator with $N_K=6,d_K=3.8,\theta_K=0.42$, (3) block generator with $N_B=11,m_B=1,M_B=14$, (4) percolation generator with $p_P=0.35,N_P=5,L_P=3.7$), and the real configuration from OSM beside each.}
    \label{fig:lhs}
\end{figure*}
%%%%%%%%%%%%%%

We can also consider the distribution of these measures within sampled urban areas. Keeping the areas with more than 10 sample points, we obtain 219 areas, for which we can compute the proportion of points within each morphological cluster. An Herfindhal diversity index on these proportions $p_k \in \left[0,1\right]$ computed as $h = 1 - \sum_k p_k^2$ ranges between 0.31 and 0.75 with an average of 0.63, suggesting very different profiles of urban areas. A chi-squared test of 10 levels of this diversity with the country is not significant (p=0.7), but the diversity index negatively correlates ($\rho = -0.11$, Fisher 95\% confidence interval $\left[-0.24,0.02\right]$) with longitude, meaning that Western cities are more diverse than Eastern cities, and more slightly with latitude ($\rho = 0.08 \left[-0.04,0.21\right]$).

@@

% TODO if no credible calib results -> discuss how it could be done. Rq : the calibration in high dimension with a projection is tricky / not much studied -- cf similar issue with diversity algorithms and the proposal of a PSE-PCA.

@@

This work can also be extended in several ways. First of all, we focused on the built environment but neglected transportation infrastructures, whereas spatial network morphogenesis models have been proposed for example by \cite{courtat2011mathematics} or \cite{raimbault2018multi} in a multi-modeling approach.
Taking into account multiple dimensions of the urban system is an important extension and hybrid models such as co-evolution models \citep{raimbault2014hybrid} should be investigated.

@@

Results obtained in this paper were computed on the vo.complex-system.eu virtual organization of the European Grid Infrastructure ( http://www.egi.eu ). We thank the European Grid Infrastructure and its supporting National Grid Initiatives (France-Grilles in particular) for providing the technical support and infrastructure.

>>>>

\usepackage[final]{changes}
%\usepackage{changes}
\usepackage{soul}
\setaddedmarkup{\textcolor{black}{\hl{#1}}}
\setdeletedmarkup{\textcolor{red}{\sout{#1}}}

@@

%
%----------------------- REVIEW 1 ---------------------
%SUBMISSION: 106
%TITLE: Generating urban morphologies at large scales
%AUTHORS: Juste Raimbault and Julien Perret
%
%----------- Overall evaluation -----------
%SCORE: 1 (weak accept)
%----- TEXT:
%Urban morphology is an interesting topic as historical biological morphology has been for the ALife community. In this manuscript the authors present a generative methodology of urban morphology and compare their results to existing European cities morphology. Their methods and results are sound and interesting and worth to be further discussed. 
%
%My main observations to this article are three --the first two interrelated:
%
%1) I think it would be interesting if the authors could better justify theoretically, why this is an interesting topic for the ALife community --beyond the use of simulations or synthetic generators which could be more in the area of Urban design? 
%
%2) Although this is an article intended to introduce the synthetic generator of urban morphologies, I guess that it would be interesting to discuss the hypothesis that inspire the generator. In other words, what are the theoretical support and hypothesis, for example, what is the relationship between self-organisation and deliberate urban design, just to mention one topic. 
%
%3) In the abstract and in the introduction it is mentioned that the topic of urban morphology may can or may have an impact on urban planning and sustainability, but these implications are completely forgotten later in the text.
%
%4) How our understanding of urban morphology would change if we could consider volume? Is that feasible now?
%
%
%
%%----------------------- REVIEW 2 ---------------------
%%SUBMISSION: 106
%%TITLE: Generating urban morphologies at large scales
%AUTHORS: Juste Raimbault and Julien Perret
%
%----------- Overall evaluation -----------
%SCORE: 2 (accept)
%----- TEXT:
%The paper is well explained and motivated, with a sound methodology.

@@

The study of forms of the built environment, and more precisely of the urban environment, has been the subject of different disciplines such as architecture, urban planning, or geography, with different approaches corresponding to various scales and processes \citep{moudon1997urban,gauthier2006mapping,kropf2009aspects}. \added{Artificial life approaches have in that context contributed to the study of generative urban processes, within the broader theoretical framework of morphogenesis }\citep{doursat2012morphogenetic}.
Establishing typologies of urban morphologies, and understanding their link with underlying urban growth processes, is nowadays a crucial issue for sustainability as a large majority of the world population live in cities and energy consumption is closely related to urban form through e.g. mobility patterns and automobile dependance \citep{newman2000sustainable}.

Although there is neither a unified definition of urban form, nor unified generative models and quantitative indicators to measure it, several approaches are close to the spirit of artificial life and generative social science~\citep{bonabeau1997classical,epstein1999agent}.
Procedural modeling \citep{watson2008procedural} aims at generating realistic cities, but is mostly focused on the visual impression given and does not consider realistic generative processes.
It is furthermore developed largely at larger scales than the one of the district \citep{Parish:2001:PMC:383259.383292}. \cite{merrell2010computer} generate in that context plans for interior of buildings, whereas \cite{cruz2017generation} use a cellular automaton model for building morphogenesis.
Approaches linked to urban planning have focused on the spatial distribution of land-use, at multiple resolutions \citep{liu2017future}, and proposed cellular automata models for urban sprawl, generally at the scale of the metropolitan area \citep{herold2003spatiotemporal}. For example, \cite{horner2007multi} proposes a link between urban form based on land-use and commuting.

@@

The quantification of urban form is in itself covered by a vast literature.
Recent work have proposed to apply deep learning techniques directly on vector data, such as \cite{2017arXiv170902939M} does, for a worldwide classification of road networks. Such an approach avoids the question of isolating relevant features.
However, as we aim at calibrating generative models, our quantification will make more sense with interpretable measures.
\cite{boeing2018measuring} proposes an extensive review of existing measures from a large extent of disciplines, their implications for planning and design, and the relation with urban complexity. \cite{webster1995urban} uses image processing techniques such as contrast or Fourier analysis, to extract synthetic descriptions of urban areas from satellite imaging. \cite{fumega2014identification} provide a typology of cities in relation with energy consumption in the perspective of climate change. \cite{rode2014cities} relate indicators of urban form with residential heat-energy demand. Other complexity-related approaches such as fractal dimensions have been introduced as for example by \cite{batty1987fractal}.

%  \cite{oliveira2016morpho} \added{ include the potential links between urban form and function in quantitative measures.}

@@

We detail now the generative models introduced. \added{The models developed below capture both bottom-up self-organizing processes and top-down planning processes. They also include different urban aspects, either filling the space with built artefacts or focusing on linking empty spaces.} A null model is also considered, to ensure the relevance of the morphological measures and the adjustments on these. It consists in a random grid generator, where each cell is occupied if a random uniform number between 0 and 1 is below a density parameter $d_R$.

%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{blocks}
        \caption{Blocks}
        \label{fig:blocks}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{expMixture}
        \caption{Kernel mixture}
        \label{fig:mixture}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{percolation}
        \caption{Network percolation}
        \label{fig:percolation}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{random}
        \caption{Random}
        \label{fig:random}
    \end{subfigure}
    \caption{Examples of patterns produced by the synthetic generators.}
    %  should provide parameters here
    \label{fig:generators}
\end{figure}
%%%%%%%%%%%%%%

@@

Simulation results and real measures are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/LGK0US}. Source code is available on the git repository of the project at \url{https://github.com/openmole/spatialdata}.
%  split repositories
The model and indicators were coded in scala langage for performance purposes. This furthermore allows a seamless integration into the OpenMOLE workflow engine for model exploration \citep{reuillon2013openmole}, which provides methods for numerical experiments (in our case sampling methods) and transparent access to high performance computation environments (model simulation were run on the European Grid Infrastructure for an equivalent of one year and one month CPU time).

@@

We compute the morphological indicators given above on a large sample on real urban areas. For practical computational reasons, we restrain our geographical area of study to European functional urban areas as provided by \cite{bretagnolle2019following}. We expect to already have a good representativity, although not universal, of existing urban forms with this sample, as it is known that European cities already have a significant morphological diversity \citep{le2015forme}. We collect building layouts from OpenStreetMap (illustrated in Figure~\ref{fig:osm}), as this source has been shown to have a good quality especially in Europe \citep{mooney2010towards}. Using the \texttt{osmosis} tool, buildings are filtered from the openstreetmap raw dump for Europe (downloaded from \url{http://download.geofabrik.de/} in March 2019) and inserted into a Postgis database, which can then be efficiently queried for a specific bounding box. Indeed, although the library developed provides a direct access to the OpenStreetMap API, query limitations do not allow such a systematic sampling.
%Our tool also provide a direct access to the classic and overpass API % not necessary ?
We sample $N=72,000$ points into polygons corresponding to urban areas, first by selecting the area with a uniform selection weighted by population of areas, then by drawing uniform spatial coordinates within the polygon with a polygon sampling heuristic. %  clarify the polygon sampler

@@

%%%%%%%%%%%%%%
\begin{figure*}
\vspace{-1.8cm}
    \centering
    (1)
    \framebox{\includegraphics[width=0.18\linewidth]{closest1_blocks}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest1_real}}\hspace{0.4cm}
    (2)
    \framebox{\includegraphics[width=0.18\linewidth]{closest2_expMixture}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest2_real}}\\\vspace{0.2cm}
    (3)
    \framebox{\includegraphics[width=0.18\linewidth]{closest3_blocks}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest3_real}}\hspace{0.4cm}
    (4)
    \framebox{\includegraphics[width=0.18\linewidth]{closest4_percolation}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest4_real}}\\
    \includegraphics[width=\linewidth]{lhscalib_projrealpcs.jpg}
    \caption{Comparison of real morphologies with patterns produced by the synthetic generators. All points are projected into the first two components computed on the real measures only. Points in dark blue correspond to real configurations. The type of generator is given by the color in legend. Dark circle correspond to the centroids of clusters on which the calibration is done, their number giving the corresponding configurations shown above. For each, we successively show the best synthetic configuration ((1) block generator with $N_B=12,m_B=5,M_B=12$, (2) exponential mixture generator with $N_K=6,d_K=3.8,\theta_K=0.42$, (3) block generator with $N_B=11,m_B=1,M_B=14$, (4) percolation generator with $p_P=0.35,N_P=5,L_P=3.7$), and the real configuration from OSM beside each.}
    \label{fig:lhs}
\end{figure*}
%%%%%%%%%%%%%%

We can also consider the distribution of these measures within sampled urban areas. Keeping the areas with more than 10 sample points, we obtain 219 areas, for which we can compute the proportion of points within each morphological cluster. An Herfindhal diversity index on these proportions $p_k \in \left[0,1\right]$ computed as $h = 1 - \sum_k p_k^2$ ranges between 0.31 and 0.75 with an average of 0.63, suggesting very different profiles of urban areas. A chi-squared test between the country and a discretization with 10 levels of this diversity is not significant (p=0.7), but the diversity index negatively correlates ($\rho = -0.11$, Fisher 95\% confidence interval $\left[-0.24,0.02\right]$) with longitude, meaning that Western cities are more diverse than Eastern cities, and more slightly with latitude ($\rho = 0.08 \left[-0.04,0.21\right]$).

@@

%  if no credible calib results -> discuss how it could be done. Rq : the calibration in high dimension with a projection is tricky / not much studied -- cf similar issue with diversity algorithms and the proposal of a PSE-PCA.

@@

This work can also be extended in several ways. First of all, we focused on the built environment but neglected transportation infrastructures, whereas spatial network morphogenesis models have been proposed for example by \cite{courtat2011mathematics} or \cite{raimbault2018multi} in a multi-modeling approach.
Taking into account multiple dimensions of the urban system is an important extension and hybrid models such as co-evolution models \citep{raimbault2014hybrid} should be investigated.
\added{Our approach can also be a preliminary step towards the study of urban sustainability issues, for example the relations between urban form and energy consumption }\citep{le2015forme}.\added{ Extending the generators with the third dimension, i.e. taking into account building heights as done by }\cite{brasebin2017apports}\added{, could also be an important component for the study of local energy efficiency.}

@@

Results obtained in this paper were computed on the vo.complex-system.eu virtual organization of the European Grid Infrastructure ( http://www.egi.eu ). We thank the European Grid Infrastructure and its supporting National Grid Initiatives (France-Grilles in particular) for providing the technical support and infrastructure. \added{This work is part of DynamiCity, a FUI project funded by BPI France, Auvergne-Rh{\^o}ne-Alpes region, Ile-de-France region and Lyon metropolis.}
